Predicting Absolute Holocellulose and Klason Lignin Contents for Peat Remains Challenging

### 2022-07-24 analysis/paper/001-reply-main.Rmd
Update manuscript 2.

$$
\begin{aligned}
\text{Y}_1[i] &\sim \text{beta}(\mu_{1}[i] \phi, (1-\mu_{1}[i])\phi)\\
Y_{2}[j] &\sim \text{beta}(\mu_{2}[j] \phi, (1-\mu_{2}[j])\phi)\\
\mu_{1} &= \text{logit}^{-1}\left(\alpha + \beta \text{X}\right)\\
\mu_{2} &= \text{logit}^{-1}\left(\alpha + \beta \text{X}_1\right)\\
\alpha &\sim \text{normal(0, 0.2)}\\
\beta &\sim \text{horseshoe}(zb, hs_{local}, hs_{global}, hs_{scale\_slab}^2 \cdot hs_{slab})\\
\phi &\sim \text{beta}(`r s1_data_holocellulose$phi_a`, `r s1_data_holocellulose$phi_b/s1_data_holocellulose$phi_scale`)\\
Y_{2}[j] &\sim \text{beta}(\mu_{\text{core}}[\text{index\_core}[j]] \phi_{\text{core}}[\text{index\_core}[j]], (1-\mu_{\text{core}}[\text{index\_core}[j]])\phi_{\text{core}}[\text{index\_core}[j]])\\
\mu_{\text{core}}[k] &\sim \text{beta}(\mu_{\text{pop}}[k] \phi_{\text{pop}}, (1-\mu_{\text{pop}}[k])\phi_{\text{pop}})\\
\mu_{\text{pop}}[k] &= \text{logit}^{-1}\left(\alpha_{\text{latitude}} + \beta_{\text{latitude}}\text{x}_{\text{latitude}}[k]\right)\\
\phi_{\text{core}} &\sim \text{gamma}(`r s1_data_holocellulose$phi_core_a`, `r s1_data_holocellulose$phi_core_b/s1_data_holocellulose$phi_core_scale`)\\
\phi_{\text{pop}} &\sim \text{gamma}(`r s1_data_holocellulose$phi_pop_a`, `r s1_data_holocellulose$phi_pop_b/s1_data_holocellulose$phi_pop_scale`)\\
\alpha_{\text{latitude}} &\sim \text{normal}(`r s1_data_holocellulose$a_latitude_mu`, `r s1_data_holocellulose$a_latitude_sigma`)\\
\beta_{\text{latitude}} &\sim \text{normal}(`r s1_data_holocellulose$b_latitude_mu`, `r s1_data_holocellulose$b_latitude_sigma`)\\
\end{aligned}
$$

Where $\text{Y}_1$ are the measured holocellulose contents of the training data; $\mu_{1}$ are the respective mean values modeled with the MIRS, only; $i$ is an index for each sample of the training data; $Y_{2}$ are the estimated peat holocellulose contents modeled simultaneously from the MIRS ($\mu_{2}$) and from the average average near-surface peat holocellulose contents per core ($\mu_{\text{core}}$); $j$ is an index for each sample in the peat data; $\text{index\_core}$ is a vector assigning each sample in the peat data to the respective core index; $k$ is an index for each unique core in the peat data. The average average near-surface peat holocellulose contents per core ($\mu_{\text{core}}$) are in turn modeled in dependency on the latitude.

The full model for Klason lignin is:

$$
\begin{aligned}
\text{Y}_1[i] &\sim \text{beta}(\mu_{1}[i] \phi, (1-\mu_{1}[i])\phi)\\
Y_{2}[j] &\sim \text{beta}(\mu_{2}[j] \phi, (1-\mu_{2}[j])\phi)\\
\mu_{1} &= \text{logit}^{-1}\left(\alpha + \beta \text{X}\right)\\
\mu_{2} &= \text{logit}^{-1}\left(\alpha + \beta \text{X}_1\right)\\
\alpha &\sim \text{normal(0, 0.2)}\\
\beta &\sim \text{horseshoe}(zb, hs_{local}, hs_{global}, hs_{scale\_slab}^2 \cdot hs_{slab})\\
\phi &\sim \text{beta}(`r s1_data_klason_lignin$phi_a`, `r s1_data_klason_lignin$phi_b/s1_data_klason_lignin$phi_scale`)\\
Y_{2}[j] &\sim \text{beta}(\mu_{\text{core}}[\text{index\_core}[j]] \phi_{\text{core}}[\text{index\_core}[j]], (1-\mu_{\text{core}}[\text{index\_core}[j]])\phi_{\text{core}}[\text{index\_core}[j]])\\
\mu_{\text{core}}[k] &\sim \text{beta}(\mu_{\text{pop}}[k] \phi_{\text{pop}}, (1-\mu_{\text{pop}}[k])\phi_{\text{pop}})\\
\mu_{\text{pop}}[k] &= \text{logit}^{-1}\left(\alpha_{\text{latitude}} + \beta_{\text{latitude}}\text{x}_{\text{latitude}}[k]\right)\\
\phi_{\text{core}} &\sim \text{gamma}(`r s1_data_klason_lignin$phi_core_a`, `r s1_data_klason_lignin$phi_core_b/s1_data_klason_lignin$phi_core_scale`)\\
\phi_{\text{pop}} &\sim \text{gamma}(`r s1_data_klason_lignin$phi_pop_a`, `r s1_data_klason_lignin$phi_pop_b/s1_data_klason_lignin$phi_pop_scale`)\\
\alpha_{\text{latitude}} &\sim \text{normal}(`r s1_data_klason_lignin$a_latitude_mu`, `r s1_data_klason_lignin$a_latitude_sigma`)\\
\beta_{\text{latitude}} &\sim \text{normal}(`r s1_data_klason_lignin$b_latitude_mu`, `r s1_data_klason_lignin$b_latitude_sigma`)\\
\end{aligned}
$$

>>>>

$$
\begin{aligned}
\text{Y}_1[i] &\sim \text{beta}(\mu_{1}[i] \phi, (1-\mu_{1}[i])\phi)\\
Y_{2}[j] &\sim \text{beta}(\mu_{2}[j] \phi, (1-\mu_{2}[j])\phi)\\
\mu_{1} &= \text{logit}^{-1}\left(\alpha + \beta \text{X}_1\right)\\
\mu_{2} &= \text{logit}^{-1}\left(\alpha + \beta \text{X}_2\right)\\
\alpha &\sim \text{normal(0, 0.2)}\\
\beta &\sim \text{horseshoe}(zb, hs_{local}, hs_{global}, hs_{scale\_slab}^2 \cdot hs_{slab})\\
\phi &\sim \text{gamma}(`r s1_data_holocellulose$phi_a`, `r s1_data_holocellulose$phi_b/s1_data_holocellulose$phi_scale`)\\
Y_{2}[j] &\sim \text{beta}(\mu_{\text{core}}[\text{index\_core}[j]] \phi_{\text{core}}[\text{index\_core}[j]], (1-\mu_{\text{core}}[\text{index\_core}[j]])\phi_{\text{core}}[\text{index\_core}[j]])\\
\mu_{\text{core}}[k] &\sim \text{beta}(\mu_{\text{pop}}[k] \phi_{\text{pop}}, (1-\mu_{\text{pop}}[k])\phi_{\text{pop}})\\
\mu_{\text{pop}}[k] &= \text{logit}^{-1}\left(\alpha_{\text{latitude}} + \beta_{\text{latitude}}\text{x}_{\text{latitude}}[k]\right)\\
\phi_{\text{core}} &\sim \text{gamma}(`r s1_data_holocellulose$phi_core_a`, `r s1_data_holocellulose$phi_core_b/s1_data_holocellulose$phi_core_scale`)\\
\phi_{\text{pop}} &\sim \text{gamma}(`r s1_data_holocellulose$phi_pop_a`, `r s1_data_holocellulose$phi_pop_b/s1_data_holocellulose$phi_pop_scale`)\\
\alpha_{\text{latitude}} &\sim \text{normal}(`r s1_data_holocellulose$a_latitude_mu`, `r s1_data_holocellulose$a_latitude_sigma`)\\
\beta_{\text{latitude}} &\sim \text{normal}(`r s1_data_holocellulose$b_latitude_mu`, `r s1_data_holocellulose$b_latitude_sigma`)\\
\end{aligned}
$$

All variables and parameters with subscript "1" refer to the training data for the MIRS prediction model (diverse organic matter samples, including leave, wood, and paper products):

- $\text{Y}_1$ are the measured holocellulose contents of the training data.
- $\mu_{1}$ are the respective average values modeled with MIRS.
- $\text{X}_1$ is a matrix containing the MIRS variables (variables in columns and samples in rows) for the training data. 
- $i$ is an index for each sample of the training data.
- $\phi$ is the precision parameter for the Beta distribution with which values of $\text{Y}_1$ are modeled.
- $\alpha$ and $\beta$ are intercept and regression coefficients for the MIRS prediction model. The $\beta$s are modeled via a regularized horseshoe prior parameterized following @Piironen.2017b as described in @Teickner.2022a.

All variables and parameters with subscript "2" refer to the peat samples on which the MIRS prediciton model is applied to describe the latitudinal gradient:

- $Y_{2}$ are the estimated peat holocellulose contents modeled simultaneously from the MIRS (via $\mu_{2}$ --- this is the same model as for $\mu_{1}$) and from the average average near-surface peat holocellulose contents per core ($\mu_{\text{core}}$). 
- $\text{X}_2$ is a matrix containing the MIRS variables (variables in columns and samples in rows) for the peat samples. 
- $\mu_{\text{core}}$ are average average near-surface peat holocellulose contents per core and is assumed to follow a Beta distribution which average value is modeled in dependency of latitude.
- $\mu_{\text{pop}}$ are average near-surface peat holocellulose contents computed for each latitude.
- $j$ is an index for each sample in the peat data. 
- $\text{index\_core}$ is a vector assigning each sample in the peat data to the respective core index. 
- $k$ is an index for each unique core in the peat data. 
- $\alpha_{\text{latitude}}$ and $\beta_{\text{latitude}}$ are intercept and regression coefficient of the linear model with which the latitudinal gradient in average near-surface peat holocellulose contents are described.
- $x_{\text{latitude}}[k]$ is the latitude for core $k$.
- $\phi_{\text{pop}}$ and $\phi_{\text{core}}$ are precision parameters for the gamma distributions for the average core near surface average holocellulose content and the core near surface average holocellulose contents, respectively.

The full model for Klason lignin is analogous to that for holocellulose:

$$
\begin{aligned}
\text{Y}_1[i] &\sim \text{beta}(\mu_{1}[i] \phi, (1-\mu_{1}[i])\phi)\\
Y_{2}[j] &\sim \text{beta}(\mu_{2}[j] \phi, (1-\mu_{2}[j])\phi)\\
\mu_{1} &= \text{logit}^{-1}\left(\alpha + \beta \text{X}\right)\\
\mu_{2} &= \text{logit}^{-1}\left(\alpha + \beta \text{X}_1\right)\\
\alpha &\sim \text{normal(0, 0.2)}\\
\beta &\sim \text{horseshoe}(zb, hs_{local}, hs_{global}, hs_{scale\_slab}^2 \cdot hs_{slab})\\
\phi &\sim \text{gamma}(`r s1_data_klason_lignin$phi_a`, `r s1_data_klason_lignin$phi_b/s1_data_klason_lignin$phi_scale`)\\
Y_{2}[j] &\sim \text{beta}(\mu_{\text{core}}[\text{index\_core}[j]] \phi_{\text{core}}[\text{index\_core}[j]], (1-\mu_{\text{core}}[\text{index\_core}[j]])\phi_{\text{core}}[\text{index\_core}[j]])\\
\mu_{\text{core}}[k] &\sim \text{beta}(\mu_{\text{pop}}[k] \phi_{\text{pop}}, (1-\mu_{\text{pop}}[k])\phi_{\text{pop}})\\
\mu_{\text{pop}}[k] &= \text{logit}^{-1}\left(\alpha_{\text{latitude}} + \beta_{\text{latitude}}\text{x}_{\text{latitude}}[k]\right)\\
\phi_{\text{core}} &\sim \text{gamma}(`r s1_data_klason_lignin$phi_core_a`, `r s1_data_klason_lignin$phi_core_b/s1_data_klason_lignin$phi_core_scale`)\\
\phi_{\text{pop}} &\sim \text{gamma}(`r s1_data_klason_lignin$phi_pop_a`, `r s1_data_klason_lignin$phi_pop_b/s1_data_klason_lignin$phi_pop_scale`)\\
\alpha_{\text{latitude}} &\sim \text{normal}(`r s1_data_klason_lignin$a_latitude_mu`, `r s1_data_klason_lignin$a_latitude_sigma`)\\
\beta_{\text{latitude}} &\sim \text{normal}(`r s1_data_klason_lignin$b_latitude_mu`, `r s1_data_klason_lignin$b_latitude_sigma`)\\
\end{aligned}
$$

### 2022-07-24 analysis/paper/002-reply-main.Rmd
Update manuscript 2.

1. Peat Klason lignin contents and relative accumulation during decomposition probably are underestimated by @Hodgkins.2018 which may have important consequences for decomposition of high-latitude peat deposits due to climate change.

2. The inaccuracies imply that the predicted values are only very rough estimates for holocellulose and Klason lignin contents and at least partly biased. An interpretation of the values given in @Hodgkins.2018 in terms of quantitative concepts for organic matter chemistry (i.e. holocellulose and Klason lignin) therefore is not possible.

Instead, we assume that if the study is reinterpreted in terms of qualitative concepts for carbohydrates and aromatics as used in the qualitative interpretation of mid infrared spectra for peat, there still is evidence for general latitudinal patterns in peat recalcitrance. However, interpretation and generalization of these findings are less straightforward. Moreover, we suggest improvements for uncertainty propagation for the models for latitudinal and temperature gradients. Remaining open questions to this approach should be addressed in future studies.

@@

In a recent evaluation of the models to predict peat holocellulose and Klason lignin contents presented in @Hodgkins.2018, we found that the training data used for the models (leaves, paper products) are not representative for peat [@Teickner.2022a]. The peat samples are more decomposed than the training data samples. Decomposition typically causes distinct features in mid infrared spectra which have complex effects on the applicability of spectral prediction models [@Teickner.2022a]. Since these features are not present in the training data, predictions of the original model are extrapolations. Following our re-evaluation, this causes large uncertainties in model predictions for both holocellulose and Klason lignin, and a comparatively large bias towards smaller Klason lignin contents, especially for more decomposed peat with Klason lignin contents >~30 mass-% [@Teickner.2022a]. Because of these uncertainties and biases, the analyses of @Hodgkins.2018 cannot be interpreted in terms of absolute holocellulose and Klason lignin contents.  
Instead, we suggest to interpret differences in peat chemistry presented by @Hodgkins.2018 not in terms of specific organic matter fractions (i.e. holocellulose and Klason lignin), but rather in terms of more qualitative concepts typically used in studies interpreting mid infrared spectra of peat (e.g. @Artz.2008; @Broder.2012; @Tfaily.2014; @Biester.2014), i.e. as "carbohydrates" and "aromatics". We assume that under such an interpretation, the conclusions of @Hodgkins.2018 probably are valid.  
The proposed reinterpretation is possible if values given in @Hodgkins.2018 are only interpreted in terms of the normalized peak heights used in the original prediction models, and not as absolute holocellulose and Klason lignin contents. Each model is a linear model with one independent variable (the normalized height of one peak); therefore a direct reinterpretation of the predicted absolute contents in terms of normalized peak heights is possible if the figure axes are relabeled as "Relative content of carbohydrates/aromatics" instead of "Estimated % carbohydrates/aromatics" (implying an absolute content of holocellulose and Klason lignin, respectively) [@Hodgkins.2018].   
An important open question to this approach of interpreting large-scale gradients in peat chemistry based on mid infrared spectra should be addressed in future studies: Are differences in specific mid infrared spectra peaks reliable indicators for differences in a specific concept for peat chemistry? For example, are differences in the peak around 1040 cm$^{-1}$ indicative for differences in carbohydrate contents (also across larger scales)?  
A similar problem arose in the discussion on the use of the C/N ratio as a good indicator for peat mass loss. This may only hold true if there have been no or negligible shifts in the vegetation composition which causes differences in the C/N ratio already in initial plant litter (e.g. @Biester.2014). For mid infrared spectra, such knowledge still has to be acquired in future studies, especially if large environmental gradients are considered as in @Hodgkins.2018.

@@

Besides a reinterpretation in terms of qualitative concepts for organic matter recalcitrance, we also believe that a more comprehensive quantification and propagation of uncertainties is important to further scrutinize the results of @Hodgkins.2018. In the original analysis, the following uncertainties are omitted: (1) As described above, the training dataset is not representative for peat and therefore predictions are biased. This uncertainty is not considered in the analyses of @Hodgkins.2018. (2) The second source of uncertainty which was omitted is the prediction uncertainty of the models used to predict holocellulose and Klason lignin contents. (3) A third source of uncertainty not considered yet is that in computing the regression model to analyze the latitudinal gradient based on average core near-surface contents, the uncertainty for these average values was not propagated.  
The first two sources of uncertainty can be omitted if only normalized peak heights (as qualitative indicators) are used instead of predicted holocellulose and Klason lignin contents, as described above. But the third source of uncertainty should also be considered when reinterpreting the results in terms of area normalized peak heights.  
To support our concerns, we reanalyzed the latitudinal gradient for holocellulose and Klason lignin, respectively, using the modified models described in @Teickner.2022a and fully considering the last two sources of uncertainty using a Bayesian approach (described in more detail in the supplementary information). With the modified models, the estimated slopes for the latitudinal gradient are more uncertain (95% confidence intervals are [`r s1_res_holocellulose %>% dplyr::pull(b_latitude) %>% quantile(probs = c(0.025, 0.975)) %>% round(1) %>% paste0(collapse = ",")`] and [`r s1_res_klason_lignin %>% dplyr::pull(b_latitude) %>% quantile(probs = c(0.025, 0.975)) %>% round(1) %>% paste0(collapse = ",")`] for holocellulose and Klason lignin, respectively; figure \@ref(fig:reply-res-p-sites-latitude) and supporting figure 3<!---\@ref(fig:reply-res-p-latitude-poc-slope)--->). A similar reanalysis of the depth profiles between the two latitude categories shows that differences are also more uncertain (figure \@ref(fig:reply-res-p-depth-profiles-latitude)). Our analysis therefore indicates larger uncertainties both in the depth profiles and the latitudinal gradient. Considering these uncertainties is important when peat properties predicted with a model are analyzed.
Again, the values presented here are only tentative because the modified models probably are also biased.

@@

Notwithstanding our critical remarks, the idea of @Hodgkins.2018 to relate mid infrared spectra to precisely quantifiable and defined organic matter fractions is important. Such fractions are relatively well understood and therefore allow us to better deduce what effects climate shifts can have on peat decomposition and recalcitrance. It is much more informative if we are able to precisely estimate a change in Klason lignin contents of e.g. 5 mass-% if average annual temperatures increase by 1$^{\circ}$C than if we are only able to make qualitative statements that the relative amount of aromatics (without describing in detail what this fraction means) increases, as is currently the case.  
When interpreting mid infrared spectra, distinguishing between precisely quantifiable concepts for organic matter chemistry, such as holocellulose and Klason lignin contents, and qualitative concepts which currently are used in interpreting peat MIRS, such as "carbohydrates and aromatics" (without specifying how this fraction could actually be measured), is crucial: Models relying on quantitative concepts are more testable and allow to actually quantify fractional changes in organic matter chemistry, whereas models built on qualitative concepts can only be refuted if there are blatant discrepancies relative to expectations (i.e. a decrease in "carbohydrates" with latitude). Ultimately, we need more precise models for peat variations in chemistry than can be developed based on qualitative concepts.  
We suggest the following steps to gain deeper insights into latitudinal gradient in peat chemistry: As outlined above, we suggest to interpret the results of @Hodgkins.2018 in terms of qualitative concepts for peat chemistry, i.e. "carbohydrates" and "aromatics". These suggestions also apply to a follow-up study using the same models [@Verbeke.2022]. Further studies are needed to check if specific peaks are good indicators for these qualitative concepts for peat chemistry, how the relation may be confounded e.g. by vegetation composition or other organic matter fractions, as could be the case especially for peaks related to aromatics [@Teickner.2022a]. This could be done by computing models to reliably predict holocellulose and Klason lignin contents of peat from mid infrared spectra and analyzing which spectral variables are relevant for this purpose and if there are confounding variables. Such models will also help us to get more precise insights into the degree to which high latitude peat accumulation may be decreased by decomposition due to rising temperatures, because they allow replacing more costly measurements. Finally, future studies should carefully consider uncertainties and fully propagate these into model estimates. Predicting absolute peat holocellulose and Klason lignin contents from infrared spectra remains challenging.

>>>>

1. Peat Klason lignin contents and relative accumulation during decomposition probably are underestimated by @Hodgkins.2018. This may have important consequences for decomposition of high-latitude peat deposits due to climate change.

2. The inaccuracies imply that the predicted values are only very rough estimates for holocellulose and Klason lignin contents and at least partly biased. An interpretation of the values given in @Hodgkins.2018 in terms of quantitative concepts for organic matter chemistry (i.e. holocellulose and Klason lignin) therefore is not possible.  
Instead, we assume that if the study is reinterpreted in terms of qualitative concepts for carbohydrates and aromatics as used in the qualitative interpretation of mid infrared spectra for peat, there still is evidence for general latitudinal patterns in peat recalcitrance. However, interpretation and generalization of these findings are less straightforward.

Moreover, we suggest improvements for uncertainty propagation for the models. Remaining open questions to this approach of analyzing global peat chemistry patterns should be addressed in future studies.

@@

In a recent evaluation of the models to predict peat holocellulose and Klason lignin contents presented in @Hodgkins.2018, we found that the training data used for the models (leaves, paper products) are not representative for peat [@Teickner.2022a].  
The peat samples are more decomposed than the training data samples. Decomposition typically causes distinct features in mid infrared spectra which have complex effects on the applicability of spectral prediction models [@Teickner.2022a]. Since these features are not present in the training data, predictions of the original models are extrapolations. Our re-evaluation showed that this causes large uncertainties in model predictions for both holocellulose and Klason lignin, and a comparatively large bias towards smaller Klason lignin contents, especially for more decomposed peat with Klason lignin contents >30 mass-% [@Teickner.2022a]. Because of these uncertainties and biases, the analyses of @Hodgkins.2018 cannot be interpreted in terms of absolute holocellulose and Klason lignin contents.  
Instead, we suggest to interpret differences in peat chemistry presented by @Hodgkins.2018 not in terms of the quantitative concepts (i.e. holocellulose and Klason lignin), but rather in terms of more qualitative concepts typically used in studies interpreting mid infrared spectra of peat: as "carbohydrates" and "aromatics" (e.g. @Artz.2008; @Broder.2012; @Tfaily.2014; @Biester.2014). We assume that under such an interpretation, the conclusions of @Hodgkins.2018 probably are valid.  
The proposed reinterpretation is possible if values given in @Hodgkins.2018 are only interpreted in terms of the normalized peak heights used in the original prediction models, and not as absolute holocellulose and Klason lignin contents. Each model is a linear model with one independent variable (the normalized height of one peak); therefore a direct reinterpretation of the predicted absolute contents in terms of normalized peak heights is possible if the figure axes are relabeled as "Relative content of carbohydrates/aromatics" instead of "Estimated % carbohydrates/aromatics" (implying an absolute content of holocellulose and Klason lignin, respectively) [@Hodgkins.2018].   
An important open question to this approach of interpreting large-scale gradients in peat chemistry based on mid infrared spectra should be addressed in future studies: Are differences in specific mid infrared spectra peaks reliable indicators for differences in a specific concept for peat chemistry? For example, are differences in the peak around 1040 cm$^{-1}$ indicative for differences in carbohydrate contents (also across larger scales)?  
A similar problem arose in the discussion on the use of the C/N ratio as a good indicator for peat mass loss. This may only hold true if there have been no or negligible shifts in the vegetation composition which causes differences in the C/N ratio already in initial plant litter (e.g. @Biester.2014). For mid infrared spectra, such knowledge still has to be acquired in future studies, especially if large environmental gradients are considered as in @Hodgkins.2018.

@@

Besides a reinterpretation in terms of qualitative concepts for organic matter recalcitrance, we also suggest that a more comprehensive quantification and propagation of uncertainties is important to further scrutinize the results of @Hodgkins.2018 because these uncertainties are not negligibly small.  
In the original analysis, the following uncertainties are omitted: (1) As described above, the training dataset is not representative for peat and therefore predictions are biased. (2) The prediction uncertainty of the models used to predict holocellulose and Klason lignin contents. (3) In computing the regression model to analyze the latitudinal gradient based on average core near-surface contents, the uncertainty for these average values was not propagated.  
The first two sources of uncertainty could be omitted if only normalized peak heights (as qualitative indicators) are used instead of predicted holocellulose and Klason lignin contents, as described above. But the third source of uncertainty should also be considered when reinterpreting the results in terms of area normalized peak heights.  
To support our concerns, we reanalyzed the latitudinal gradient for holocellulose and Klason lignin, respectively, using the modified models described in @Teickner.2022a and fully considering the last two sources of uncertainty using a Bayesian approach (described in more detail in the supplementary information). With the modified models, the estimated slopes for the latitudinal gradient are more uncertain (95% confidence intervals are [`r s1_res_holocellulose %>% dplyr::pull(b_latitude) %>% quantile(probs = c(0.025, 0.975)) %>% round(1) %>% paste0(collapse = ",")`] and [`r s1_res_klason_lignin %>% dplyr::pull(b_latitude) %>% quantile(probs = c(0.025, 0.975)) %>% round(1) %>% paste0(collapse = ",")`] for holocellulose and Klason lignin, respectively; figure \@ref(fig:reply-res-p-sites-latitude) and supporting figure 3<!---\@ref(fig:reply-res-p-latitude-poc-slope)--->). A similar reanalysis of the depth profiles between the two latitude categories shows that differences in depth patterns are also more uncertain (figure \@ref(fig:reply-res-p-depth-profiles-latitude)).  
Our analysis therefore indicates larger uncertainties both in the depth profiles and the latitudinal gradient. Considering these uncertainties is important when peat properties predicted with a model are analyzed.
Again, the values presented here are only tentative because the modified models probably are also biased.

@@

Notwithstanding our critical remarks, the idea of @Hodgkins.2018 to relate mid infrared spectra to precisely quantifiable and defined organic matter fractions is important. Such fractions are relatively well understood and therefore allow us to better deduce what effects climate shifts can have on peat decomposition and recalcitrance: It is much more informative if we are able to precisely estimate a change in Klason lignin contents of e.g. 5 mass-% if average annual temperatures increase by 1$^{\circ}$C than if we are only able to make qualitative statements that the relative amount of aromatics (without describing in detail what this fraction means) increases, as is currently the case.  
When interpreting mid infrared spectra, distinguishing between precisely quantifiable concepts for organic matter chemistry, such as holocellulose and Klason lignin contents, and qualitative concepts which currently are used in interpreting peat MIRS, such as "carbohydrates and aromatics" (without specifying how this fraction could actually be measured), is crucial: Models relying on quantitative concepts are more testable and allow to actually quantify fractional changes in organic matter chemistry, whereas models built on qualitative concepts can only be refuted if there are blatant discrepancies relative to expectations (i.e. a decrease in "carbohydrates" with latitude). Ultimately, we need more precise models for variations in peat chemistry which can be developed based on qualitative concepts.  
We suggest the following steps to gain deeper insights into latitudinal gradient in peat chemistry: As outlined above, we suggest to interpret the results of @Hodgkins.2018 in terms of qualitative concepts for peat chemistry, i.e. "carbohydrates" and "aromatics". These suggestions also apply to a follow-up study using the same models [@Verbeke.2022]. Further studies are needed to check if specific peaks are good indicators for these qualitative concepts for peat chemistry, how the relation may be confounded e.g. by vegetation composition or other organic matter fractions, as could be the case especially for peaks related to aromatics [@Teickner.2022a]. This could be done by computing models to reliably predict holocellulose and Klason lignin contents of peat from mid infrared spectra and analyzing which spectral variables are relevant for this purpose and if there are confounding variables. Such models will also help us to get more precise insights into the degree to which high latitude peat accumulation may be decreased by decomposition due to rising temperatures, because they allow replacing more costly measurements. Finally, future studies should carefully consider uncertainties and fully propagate these into model estimates if it is unknown whether such uncertainties are negligible. Predicting absolute peat holocellulose and Klason lignin contents from infrared spectra remains challenging.
